visitation_bias

### 2021-06-15 code/ms2_data.Rmd
almost there...

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi too rare
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  mutate(visited = rep(1, n())) %>% # binary visitation
  select(site, date, bb.sp, plant.sp, visited) %>%
  left_join(fl_traits) %>%
  select(site, date, bb.sp, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
  unite(sample.id, site, date) %>%
  mutate(sample.id = factor(sample.id),
         bb.sp = factor(bb.sp), 
         ktype = factor(ktype),
         visited = as.integer(visited))

>>>>

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi too rare
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  mutate(visited = rep(1, n())) %>% # binary visitation
  select(site, date, bb.sp, caste, plant.sp, visited) %>%
  left_join(fl_traits) %>%
  select(site, date, bb.sp, caste, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
  unite(sample.id, site, date) %>%
  mutate(sample.id = factor(sample.id),
         bb.sp = factor(bb.sp), 
         ktype = factor(ktype),
         visited = as.integer(visited))

### 2021-06-15 code/ms2_nullmodels.Rmd
almost there...

@@

@@

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

@@

ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
```

@@

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", null_model_overall, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)
```

@@

gam1_sum <- summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)
```

@@

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
```

>>>>

visitation_workers <- visitation %>%
  filter(caste == "worker")

@@

vis_spp <- visitation %>%
  left_join(fl_traits) %>%
  group_by(sample.id, bb.sp, plant.genus, ktype) %>%
  summarize(visits = n()) %>%
  group_by(bb.sp, sample.id) %>%
  mutate(prop.visits = visits/sum(visits))

view(filter(vis_spp, ktype == "1"))

ggplot(filter(vis_spp, ))

@@

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()

################################
### Repeat, excluding queens ###
################################

# The "consumers" table input to generate_null_net()
con_workers <- visitation_workers %>%
  #filter(visited == 1) %>%
  mutate(individual.id = row_number()) %>%
  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  arrange(sample.id) %>%
  ungroup() %>%
  as.data.frame()

# The "resources" table input to generate_null_net()
res_workers <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
  ungroup() %>%
  droplevels() %>%
  as.data.frame() %>%
  semi_join(con_workers, by = "sample.id") %>%
  arrange(sample.id)

# Generate null model
null1_workers <- generate_null_net(consumers = con_workers[, 3:12],
                           resources = res_workers[, 2:10],
                           sims = 500, # raised from default of 100 (takes a while to run)
                           data.type = "names",
                           # c.samples = con$sample.id, # parsed by sample
                           # r.samples = res$sample.id,
                           c.samples = con_workers[,1], # parsed by sample
                           r.samples = res_workers[,1],
                           summary.type = "mean") # summarizing with the mean instead of the default sum standardizes across our bumble bee species that vary in total abundance,

saveRDS(null1_workers, file = "../output/null1_workers.rds")
null1_workers <- readRDS("../output/null1_workers.rds")

# Tabulate output and generate tests
null_tab1_workers <- test_interactions(null1_workers, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

@@

ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)

####################
### Workers only ###
####################

ggplot(null_tab1_workers, aes(Observed, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~fct_rev(bb.sp)) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_spwise_preference_workers.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_spwise_preference_workers.png", height = 6, width = 7)
```

@@

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", null_model_overall, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)

####################
### Workers only ###
####################

null_model_overall_workers <- ggplot(null_tab1_workers, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Selectivity")

ggsave("../ms_2/figures/null_model_ktype_preference_workers.pdf", null_model_overall, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference_workers.png", null_model_overall, height = 6, width = 7)
```

@@

gam1_sum <- summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)

####################
### Workers only ###
####################

gam1_workers <- gam(preference ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1_workers) %>% getViz()

gam1_sum_workers <- summary(gam1_workers)
check.gamViz(gam1_workers)
#plot.gamViz(gam1_workers, allTerms = TRUE)
```

@@

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)

####################
### Workers only ###
####################

# Tabulate model summary
gam1_s_table_workers <- gam1_sum_workers$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype))

# Extract predictions
gam1_pred_workers <- predict_gam(gam1_workers) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit)) %>%
  left_join(gam1_s_table_workers) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))

# Plot
null_model_pbl_GAM_workers <- ggplot(gam1_pred_workers, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1_workers, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Selectivity", 
       color = "Species", linetype = "p < 0.05") +
  theme_light(10)

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference_workers.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference_workers.png", null_model_pbl_GAM, height = 6, width = 7)
```
